---
title: "EDA LA1"
author: "Kishore, Lakshminarasimhan"
date: "2023-05-16"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

In this part, we are loading and reading
a CSV file named “train.csv” from the specified path and assigning it to the dataframe df. Finally,
we are viewing the dataframe df by using the head() function.
```{r}
  df<-read.csv("train.csv")
  head(df)
```

The last 6 rows are accessed using the tail() function
```{r}
  tail(df)
```
The structure of the df is shown below

```{r}
str(df)

```
the summary of the df is printed using the summary() function

```{r}
summary(df)

```
All the dimensional attributes of the df is checked below like total dimension, the number of rows, number of columns etc.

```{r}
dim(df)
```
```{r}
nrow(df)
```
```{r}
ncol(df)
```
The column names, class and type of data frame df is described below using its respective functions.

```{r}
names(df)
```
```{r}
class(df)
```
```{r}
typeof(df)
```
```{r}
is.data.frame(df)
```
```{r}
ls()
```
```{r}
list(df)
```
```{r}
#history()
```
 
```{r}
save.image("eda_la1.RData") 
```

```{r}
save(df,file="saved.RData")
```

```{r}
samp<-load("saved.RData")
samp
```
column wise accessing is done and operations performed onto the respective columns

```{r}
df$Loan_ID # Accessing a column of the data frame using $ notation

```

```{r}
df[, "Education"] # Accessing a column of the data frame using bracket notation

```

```{r}
df[1:10,] # Accessing the first 10 rows of the data frame

```

```{r}
df[df$Dependents > 1,] # Accessing rows where a certain condition is met

```

```{r}
df[df$LoanAmount %in% c(168, 112),] # Accessing rows where a certain value is present

```

```{r}
df[order(df$Loan_Amount_Term),] # Sorting the data frame by a column

```

```{r}
df[, c("Loan_Amount_Term", "LoanAmount")] # Selecting multiple columns from the data frame

```

```{r}
df$column_name <- NULL # Deleting a column from the data frame

```
```{r}
names(df)[1]
```

```{r}
names(df)[3] <- "Marital_Status" # Renaming a column in the data frame
names(df)[3]
```

```{r}
df1 <- df[, c(2, 1)] # Reordering columns in the data frame
df1
```
```{r}
df
```
The coapplicantincome column is re-coded from numerical to categorical for easy understanding

```{r}
df$cat_CoapplicantIncome <- cut(df$CoapplicantIncome, breaks = c(-1, 5000, 10000, Inf), labels = c("low", "medium", "high")) # Re-coding a continuous variable to a categorical variable
```
```{r}
df
```
The LoanAmount varible is log normalized as to reduce the amount and time of calculations that are performed onto that column

```{r}
df$new_LoanAmount <- log(df$LoanAmount) # Transforming a variable using a mathematical function
df
```
The library dplyr is imported for further use

```{r}
library(dplyr)
```
dplyr package functions are used for grouping the data and sumarising the data 

```{r}
df2 <- group_by(df, Education) # Grouping the data frame by a column
df2
```
```{r}
summarise(df, mean_ApplicantIncome = mean(ApplicantIncome)) # Summarizing data by groups
```

The next step summarises the data by groups with standard errors and confidence intervals

```{r}
summarise(df,sd_ApplicantIncome = mean(ApplicantIncome), sd = sd(ApplicantIncome)) # Summarizing data by groups with standard errors and confidence intervals
```


The min max of the respective numerical column is found out
```{r}
#?min
min(df[8])
max(df[8])
```

```{r}
min(df[7])
max(df[7])
```
```{r}
min(df[15])
max(df[15])
```
```{r}

```

Now the dataframe is arranged according to the variable of interest such as LoanAmount, LoanAmountTerm, CoapplicantIncome and Dependents
```{r}
arrange(df, LoanAmount)
```

```{r}
arrange(df, Loan_Amount_Term)
```

```{r}
arrange(df, CoapplicantIncome)
```

```{r}
arrange(df, Dependents)
```
```{r}
df
```
The following code shows standard deviation of each column with itself.

```{r}
sd(df$CoapplicantIncome)
```

```{r}
sd(df$LoanAmount)
```

```{r}
sd(df$ApplicantIncome)
```


```{r}
mean(df$Loan_Amount_Term)

```

Bar chart is plotted for the whole dataframe for easy visualization
```{r}
plot(df)
```


Histograms are plotted to understand the trend of values in the columns
```{r}
hist(df$LoanAmount)
```
```{r}
hist(df$CoapplicantIncome)
```


```{r}
hist(df$ApplicantIncome)
```
```{r}
hist(df$LoanAmount)

```

Pie chart is plotted to understand the distribution of column data with respect to the data frame
```{r}
pie(table(df$ApplicantIncome))
```

Here the bar plots are plotted against 2 different columns to find trends instead of just using the normal single column values

```{r}
plot(x=df$ApplicantIncome,y=df$CoapplicantIncome)
```
```{r}
boxplot(df$ApplicantIncome,df$CoapplicantIncome)
```

